---
title: "Prediction Assignment"
author: "RC"
date: "5 November 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation

Read in the training and test datasets.
```{r}
training <- read.csv("pml-training.csv", na.strings = c("NA", "#DIV/0!"))
testing <- read.csv("pml-testing.csv", na.strings = c("NA", "#DIV/0!"))
```

Clean the datasets, removing columns with missing data (NAs).
```{r}
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]
```

Remove the first 7 columns in each dataset as they do not contain data that can be used for predictions (X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window, num_window).
```{r}
training <- training[, -c(1:7)]
testing <- testing[, -c(1:7)]
```

Split the datasets into further training and testing sets for prediction. 
```{r message=FALSE, echo=FALSE}
set.seed(1234)
suppressWarnings(library(caret))
inTrain <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
training_new <- training[inTrain,]
testing_new <- training[-inTrain,]
```

## Model Prediction
We will try fitting a linear discriminant analysis (LDA) and gradient boosting (GBM) models.
```{r}
fit_Control <- trainControl(method="cv", number = 5, allowParallel = TRUE)

fit_LDA <- train(classe ~ ., data=training_new, method="lda", trControl = fit_Control)

fit_GBM <- train(classe ~ ., data=training_new, method="gbm", trControl = fit_Control, verbose = FALSE)
```

### Now, we test the accurary of our models on the training data and testing data.

LDA Model
```{r}
prediction_LDA <- predict(fit_LDA, training_new)
confusionMatrix(prediction_LDA, training_new$classe)$overall[1]

prediction_LDA_test <- predict(fit_LDA, testing_new)
confusionMatrix(prediction_LDA_test, testing_new$classe)$overall[1]

```

GBM Model
```{r}
prediction_GBM <- predict(fit_GBM, training_new)
confusionMatrix(prediction_GBM, training_new$classe)$overall[1]

prediction_GBM_test <- predict(fit_GBM, testing_new)
confusionMatrix(prediction_GBM_test, testing_new$classe)$overall[1]
```

We see from this that the latter GBM model has much higher accuracy.

### Lastly, we make predictions on the original testing set.

We will use the more accurage GBM model.

```{r}
predict_GBM <- predict(fit_GBM, testing)
predict_GBM
```
